\pkg{dataMaid}: your assistant for documenting supervised data quality screening in \proglang{R}

### 2017-11-14 latex/article_vol3.Rnw
Updated system calls to remove output on all platforms
Did spell checking and fixed mistakes. Note: This file has not been compiled!

  We describe an \proglang{R} package, \pkg{dataMaid}, which
  implements an extensive and customizeable suite of quality
  assessment aids that can be applied to a dataset in order to
  identify potential problems in its variables. The results are
  presented in an auto-generated, non-technical, stand-alone overview
  document intended to be perused by an investigator with an
  understanding of the variables in the data, but not necessarily
  knowledge of \proglang{R}. Thereby, \pkg{dataMaid} aids the dialogue
  between data analysts and field experts, while also providing easy
  documentation of reproducible data quality screening. Moreover, the
  \pkg{dataMaid} solution changes the data screening process from the
  usual ad hoc approach to a systematic, well-documented endeavor.
  \pkg{dataMaid} also provides a suite of more typical \proglang{R}
  tools for interactive data quality assessment and -screening, where
  the data inspections are executed directly in the \proglang{R}
  console.

@@

But even when tools are available for identifying problems in a
dataset, the activity of data cleaning still suffers from a challenge
that has recieved increasing attention in the scientitic communities
in the later years: Data cleaning is not very straight forward to
document and therefore, reproducibility suffers. We present a new
\proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most
central purpose is to facilitate a supervised data quality screening
workflow where documentation is thoroughly integrated rather than an
add-on. This is accomplished by structuring the data screening around
auto-generated data overview reports that summarize and flags
potential problems in the dataset.

@@

A number of \proglang{R} packages made for other pre-analysis steps
are already available, including \pkg{janitor} \citep{janitor},
\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
\citep{validate}, and \pkg{assertr} \citep{assertr}. These packages
focus on different stages of the pre-analysis work.  \pkg{janitor}
provides tools for data import with a particular emphasis on the
challenges of getting neat data frames from Microsoft Excel data
files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
\pkg{DataCombine} go a few steps further by providing a wide array of
extremely powerful tools for data wrangling, including a number of
particularly useful functions for merging and working with very large
datasets. When it comes to actual data cleaning, however, the options
are fewer.  \pkg{validate} (and the similar packages \pkg{editrules}
\citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
same authors) and \pkg{assertive} offer tools for identifying errors
in a dataset by checking the state of the variable given a set of
pre-specified rules, and their focus is on internal validity rather
than general data screening. In practice, this means that quite elegant
tools for, e.g., linear restraints among the variables in a dataset
can be applied, but looking for potentially miscoded missing values is
not really feasible. The main difference between these two challenges
is the direction in which the data is inspected: While linear
constraints work observation-wise with no ambiguity, determining
whether or not something is a miscoded missing value often requires
knowledge about the full variable (e.g. range or data type), and thus
it should be performed variable-wise. \pkg{validate} does not
currently allow for user-defined extensions of the latter type,
thereby limiting its data cleaning potential.  Automatic data
correction functions are also provided by \pkg{validate} which we
consider to be quite a dangerous cocktail: all power is given to the
the computer with no human supervision, and investigators are less
likely to make an active, case-specific choice regarding the handling
of the potential errors.  Finally, no tools have been made to easily
document exactly which checks and preliminary results were used in the
data cleaning process. The \pkg{assertr} package provide very similar
--- and very nice --- tools to those of \pkg{validate}, but without
any amibitions of conducting auto-cleaning.

@@

%Please note that this description of what \pkg{dataMaid} contains is
%usefully supplemented by a tutorial into how it can be extended. Every
%dataset is different, and some datasets might include problems that
%cannot be detected by our data checking functions. 
%Therefore,
\pkg{dataMaid} was designed to be easily extendend with user-supplied
functions for summarizing, viusalizing and checking data. In the package, we have
provided a vignette in which we describe how \pkg{dataMaid} extensions
can be made, such that they are integrate with the
\code{makeDataReport()} function and with the other tools available in
\pkg{dataMaid}.

@@

The \code{makeDataReport()} function is the primary workhorse of
\pkg{dataMaid} and it is the only function needed to generate a data
report using the standard battery of tests. The data report itself is
an overview document, intended for reading by humans, in either pdf,
html or word (.docx) format. Appendix \ref{sec:appendix1} provides an
example of a data report, produced by calling \code{makeDataReport()}
on the dataset \code{toyData} available in \pkg{dataMaid}. The first
two pages (excluding the frontpage) of this data report are shown in
Figure~\ref{fig:example1} and the following two pages are shown in
Figure~\ref{fig:example2}. \code{toyData} is a very small ($15$
observations of $6$ variables), artificial dataset which was created
with a lot of potential errors to illustrate the main capabilities of
\pkg{dataMaid}. Section~\ref{sec:bigExample} shows an example of a data sreening
process with a real dataset. The following commands load the dataset
and produce the report:

@@

The arguments to \code{makeDataReport()} can be used to modify the
contents and the look of the data report according to the user's
needs. The most commonly used arguments are summarized in
Table~\ref{table.cleanFormals} and they are grouped according to the
part of the data assesment and report generation they influence. In
order to understand this distinction, a glimpse of the inner structure
of \code{makeDataReport()} is shown in
Figure~\ref{figure:cleanStructure}. Below, we present a few examples
on how to use the arguments from Table \ref{table.cleanFormals} to
influence the output of a \code{makeDataReport()} call.

@@

\begin{center}
\begin{tikzpicture}[node distance = 2cm, auto,thick,scale=0.75, every node/.style={transform shape}]
    % Place nodes
    \node [block] (init) {Get next variable and run pre-checks};
    \node [cloud, above of=init] (input) {Input \code{data.frame} or \code{tibble}};
%    \node [cloud, right of=init] (system) {system};
    \node [decision, right of=init, node distance=4cm] (precheck) {Is variable suitable for inclusion};
    \node [block, right of=precheck, node distance=4cm] (summarize)
    {Run \code{summarize()} to produce summary table};
    \node [block, below of=summarize, node distance=3cm] (visualize)
    {Run \code{visualize()} to plot variable};
    \node [block, below of=visualize, node distance=2.5cm] (check)
    {Call \code{check()} to run error checks};
    \node [decision, below of=check, node distance=2.7cm] (done) {More
      variables?};
    \node [block, right of=done, node distance=3.5cm] (stop) {Write
      \proglang{R} markdown file};
    \node [cloud, below of=stop, node distance=3.5cm] (render) {Render
      markdown and possiby open};
    % Draw edges
    \path [line] (summarize) -- (visualize);
    \path [line] (visualize) -- (check);
    \path [line] (check) -- (done);
    \path [line] (done.south) -- +(0,-10pt) -| node [near start] {yes} (init);
%    \path [line] (identify) -- (evaluate);
%    \path [line] (evaluate) -- (decide);
%    \path [line] (init) -| node [near start] {yes} (precheck);
    \path [line] (init) -- (precheck);
    \path [line] (precheck) -- node [near start] {yes} (summarize);
    \path [line] (precheck) |- node [near start] {no} (done);
    \path [line] (done) -- node [near start] {no} (stop);
%    \path [line] (update) |- (identify);
%    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (input) -- (init);
    \path [line] (stop) -- (render);
%    \path [line,dashed] (system) -- (init);
%    \path [line,dashed] (system) |- (evaluate);
\end{tikzpicture}
\end{center}

@@

 \textbf{\code{checkFunction}s} \smallskip \\
 \quad \code{identifyCaseIssues} & Identify case issues &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyLoners} & Identify levels with $<$ 6 obs. & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyMissing} & Identify miscoded missing values &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} &  \\
 \quad \code{identifyNums} & Identify misclassified numeric or integer variables & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyOutliers} & Identify outliers &  & & \blue{$\times$} & & \blue{$\times$} & \blue{$\times$} \\
 \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
 \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
 \quad \code{isSingular} & Check if the variable contains only a single value & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
 \quad \code{isKey} & Check if the variable is a key & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
 \quad \code{isSupported} & Check if the variable is among the supported variable types & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
 \hline
\end{tabular}
\caption{Overview of all summary-, visual- and  check functions currently implemented in \pkg{dataMaid}. The variable
  classes C, F, I, L, B, N, and D, refer to \code{character}, \code{factor},
  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively. The default settings of \code{makeDataReport()} are marked in blue.}
\label{table.SVCfunctions}
\end{table}

@@

Now, if we only wanted to apply the function to identify whitespace
for factor variables, then we would need provide this information for \code{setChecks()}:

@@

We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a useful overview report and how the interactive tools can subsequently be used to assist the actual data cleaning process. More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data. \code{presidentData} is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger. The dataset contains one observation per president and has the following variables:
\begin{description}
\item[\code{lastName}] The last name of the president.
\item[\code{firstName}] The first name of the president.
\item[\code{orderOfPresidency}] The number in the order of presidents.
\item[\code{birthday}] The birthday of the president.
\item[\code{stateOfBirth}] The state in which the president was born.
\item[\code{assassinationAttempt}] Was there an assassination attempt on the president?
\item[\code{sex}] The sex of the president.
\item[\code{ethnicity}] The ethnicity of the president.
\item[\code{presidencyYears}] The duration of the presidency.
\item[\code{ageAtInauguration}] The age of the president at inauguration.
\item[\code{favoriteNumber}] The favourite number of the president (fictional).
\end{description}

@@

On page 3, we see the contents for the three first variables. Here, we identifify a prefixed whitespace in the lastname entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g. Stata, and therefore, it is flagged as a potential miscoded missing value. The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.

@@

A few of the identified problems have easy fixes that need no further discussion. We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:

@@

The three problems can easily be solved by use of the arguments of \code{makeDataReport()}. Whether or not the outputted file is opened can be controlled through the argument \code{open}. How much information is printed in the console can be adjusted by using the argument \code{quiet}. And convenintely introducing small alterations of the file names can be obtained by use of the \code{vol} argument. For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:

@@

Now only the checking results are printed, and only for variables
where problems were identified. An even more minimal output  can be obtained directly in the console by using the \code{check()} function interactively. When called on a \code{data.frame}, this function produces a list (of
variables) of lists (of checks) of lists (or rather,
\code{checkResult}s). Thus, the overall problem status of each variable
can easily be unravelled using the list manipulation function
\code{sapply()}:

@@

>>>>

  We describe an \proglang{R} package, \pkg{dataMaid}, which
  implements an extensive and customizable suite of quality
  assessment aids that can be applied to a dataset in order to
  identify potential problems in its variables. The results are
  presented in an auto-generated, non-technical, stand-alone overview
  document intended to be perused by an investigator with an
  understanding of the variables in the data, but not necessarily
  knowledge of \proglang{R}. Thereby, \pkg{dataMaid} aids the dialogue
  between data analysts and field experts, while also providing easy
  documentation of reproducible data quality screening. Moreover, the
  \pkg{dataMaid} solution changes the data screening process from the
  usual ad hoc approach to a systematic, well-documented endeavor.
  \pkg{dataMaid} also provides a suite of more typical \proglang{R}
  tools for interactive data quality assessment and -screening, where
  the data inspections are executed directly in the \proglang{R}
  console.

@@

But even when tools are available for identifying problems in a
dataset, the activity of data cleaning still suffers from a challenge
that has received increasing attention in the scientific communities
in the later years: Data cleaning is not very straight forward to
document and therefore, reproducibility suffers. We present a new
\proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most
central purpose is to facilitate a supervised data quality screening
workflow where documentation is thoroughly integrated rather than an
add-on. This is accomplished by structuring the data screening around
auto-generated data overview reports that summarize and flags
potential problems in the dataset.

@@

A number of \proglang{R} packages made for other pre-analysis steps
are already available, including \pkg{janitor} \citep{janitor},
\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
\citep{validate}, and \pkg{assertr} \citep{assertr}. These packages
focus on different stages of the pre-analysis work.  \pkg{janitor}
provides tools for data import with a particular emphasis on the
challenges of getting neat data frames from Microsoft Excel data
files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
\pkg{DataCombine} go a few steps further by providing a wide array of
extremely powerful tools for data wrangling, including a number of
particularly useful functions for merging and working with very large
datasets. When it comes to actual data cleaning, however, the options
are fewer.  \pkg{validate} (and the similar packages \pkg{editrules}
\citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
same authors) and \pkg{assertive} offer tools for identifying errors
in a dataset by checking the state of the variable given a set of
pre-specified rules, and their focus is on internal validity rather
than general data screening. In practice, this means that quite elegant
tools for, e.g., linear restraints among the variables in a dataset
can be applied, but looking for potentially miscoded missing values is
not really feasible. The main difference between these two challenges
is the direction in which the data is inspected: While linear
constraints work observation-wise with no ambiguity, determining
whether or not something is a miscoded missing value often requires
knowledge about the full variable (e.g. range or data type), and thus
it should be performed variable-wise. \pkg{validate} does not
currently allow for user-defined extensions of the latter type,
thereby limiting its data cleaning potential.  Automatic data
correction functions are also provided by \pkg{validate} which we
consider to be quite a dangerous cocktail: all power is given to the
the computer with no human supervision, and investigators are less
likely to make an active, case-specific choice regarding the handling
of the potential errors.  Finally, no tools have been made to easily
document exactly which checks and preliminary results were used in the
data cleaning process. The \pkg{assertr} package provide very similar
--- and very nice --- tools to those of \pkg{validate}, but without
any ambitions of conducting auto-cleaning.

@@

%Please note that this description of what \pkg{dataMaid} contains is
%usefully supplemented by a tutorial into how it can be extended. Every
%dataset is different, and some datasets might include problems that
%cannot be detected by our data checking functions. 
%Therefore,
\pkg{dataMaid} was designed to be easily extended with user-supplied
functions for summarizing, visualizing and checking data. In the package, we have
provided a vignette in which we describe how \pkg{dataMaid} extensions
can be made, such that they are integrate with the
\code{makeDataReport()} function and with the other tools available in
\pkg{dataMaid}.

@@

The \code{makeDataReport()} function is the primary workhorse of
\pkg{dataMaid} and it is the only function needed to generate a data
report using the standard battery of tests. The data report itself is
an overview document, intended for reading by humans, in either pdf,
html or word (.docx) format. Appendix \ref{sec:appendix1} provides an
example of a data report, produced by calling \code{makeDataReport()}
on the dataset \code{toyData} available in \pkg{dataMaid}. The first
two pages (excluding the front page) of this data report are shown in
Figure~\ref{fig:example1} and the following two pages are shown in
Figure~\ref{fig:example2}. \code{toyData} is a very small ($15$
observations of $6$ variables), artificial dataset which was created
with a lot of potential errors to illustrate the main capabilities of
\pkg{dataMaid}. Section~\ref{sec:bigExample} shows an example of a data screening
process with a real dataset. The following commands load the dataset
and produce the report:

@@

The arguments to \code{makeDataReport()} can be used to modify the
contents and the look of the data report according to the user's
needs. The most commonly used arguments are summarized in
Table~\ref{table.cleanFormals} and they are grouped according to the
part of the data assessment and report generation they influence. In
order to understand this distinction, a glimpse of the inner structure
of \code{makeDataReport()} is shown in
Figure~\ref{figure:cleanStructure}. Below, we present a few examples
on how to use the arguments from Table \ref{table.cleanFormals} to
influence the output of a \code{makeDataReport()} call.

@@

\begin{center}
\begin{tikzpicture}[node distance = 2cm, auto,thick,scale=0.75, every node/.style={transform shape}]
    % Place nodes
    \node [block] (init) {Get next variable and run pre-checks};
    \node [cloud, above of=init] (input) {Input \code{data.frame} or \code{tibble}};
%    \node [cloud, right of=init] (system) {system};
    \node [decision, right of=init, node distance=4cm] (precheck) {Is variable suitable for inclusion};
    \node [block, right of=precheck, node distance=4cm] (summarize)
    {Run \code{summarize()} to produce summary table};
    \node [block, below of=summarize, node distance=3cm] (visualize)
    {Run \code{visualize()} to plot variable};
    \node [block, below of=visualize, node distance=2.5cm] (check)
    {Call \code{check()} to run error checks};
    \node [decision, below of=check, node distance=2.7cm] (done) {More
      variables?};
    \node [block, right of=done, node distance=3.5cm] (stop) {Write
      \proglang{R} markdown file};
    \node [cloud, below of=stop, node distance=3.5cm] (render) {Render
      markdown and possibly open};
    % Draw edges
    \path [line] (summarize) -- (visualize);
    \path [line] (visualize) -- (check);
    \path [line] (check) -- (done);
    \path [line] (done.south) -- +(0,-10pt) -| node [near start] {yes} (init);
%    \path [line] (identify) -- (evaluate);
%    \path [line] (evaluate) -- (decide);
%    \path [line] (init) -| node [near start] {yes} (precheck);
    \path [line] (init) -- (precheck);
    \path [line] (precheck) -- node [near start] {yes} (summarize);
    \path [line] (precheck) |- node [near start] {no} (done);
    \path [line] (done) -- node [near start] {no} (stop);
%    \path [line] (update) |- (identify);
%    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (input) -- (init);
    \path [line] (stop) -- (render);
%    \path [line,dashed] (system) -- (init);
%    \path [line,dashed] (system) |- (evaluate);
\end{tikzpicture}
\end{center}

@@

 \textbf{\code{checkFunction}s} \smallskip \\
 \quad \code{identifyCaseIssues} & Identify case issues &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyLoners} & Identify levels with $<$ 6 obs. & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyMissing} & Identify miscoded missing values &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} &  \\
 \quad \code{identifyNums} & Identify misclassified numeric or integer variables & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{identifyOutliers} & Identify outliers &  & & \blue{$\times$} & & \blue{$\times$} & \blue{$\times$} \\
 \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
 \quad \code{identifyWhitespace} & Identify prefixed and suffixed white space &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
 \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
 \quad \code{isSingular} & Check if the variable contains only a single value & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
 \quad \code{isKey} & Check if the variable is a key & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
 \quad \code{isSupported} & Check if the variable is among the supported variable types & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
 \hline
\end{tabular}
\caption{Overview of all summary-, visual- and  check functions currently implemented in \pkg{dataMaid}. The variable
  classes C, F, I, L, B, N, and D, refer to \code{character}, \code{factor},
  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively. The default settings of \code{makeDataReport()} are marked in blue.}
\label{table.SVCfunctions}
\end{table}

@@

Now, if we only wanted to apply the function to identify white space
for factor variables, then we would need provide this information for \code{setChecks()}:

@@

We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a useful overview report and how the interactive tools can subsequently be used to assist the actual data cleaning process. More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data. \code{presidentData} is a slightly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger. The dataset contains one observation per president and has the following variables:
\begin{description}
\item[\code{lastName}] The last name of the president.
\item[\code{firstName}] The first name of the president.
\item[\code{orderOfPresidency}] The number in the order of presidents.
\item[\code{birthday}] The birthday of the president.
\item[\code{stateOfBirth}] The state in which the president was born.
\item[\code{assassinationAttempt}] Was there an assassination attempt on the president?
\item[\code{sex}] The sex of the president.
\item[\code{ethnicity}] The ethnicity of the president.
\item[\code{presidencyYears}] The duration of the presidency.
\item[\code{ageAtInauguration}] The age of the president at inauguration.
\item[\code{favoriteNumber}] The favorite number of the president (fictional).
\end{description}

@@

On page 3, we see the contents for the three first variables. Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g. Stata, and therefore, it is flagged as a potential miscoded missing value. The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.

@@

A few of the identified problems have easy fixes that need no further discussion. We remove the prefixed white space from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:

@@

The three problems can easily be solved by use of the arguments of \code{makeDataReport()}. Whether or not the outputted file is opened can be controlled through the argument \code{open}. How much information is printed in the console can be adjusted by using the argument \code{quiet}. And conveniently introducing small alterations of the file names can be obtained by use of the \code{vol} argument. For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:

@@

Now only the checking results are printed, and only for variables
where problems were identified. An even more minimal output  can be obtained directly in the console by using the \code{check()} function interactively. When called on a \code{data.frame}, this function produces a list (of
variables) of lists (of checks) of lists (or rather,
\code{checkResult}s). Thus, the overall problem status of each variable
can easily be unraveled using the list manipulation function
\code{sapply()}:

@@

\section*{Acknowledgements}
\label{acknowledgements}
This work was supported by The Lundbeck Foundation, Trygfonden and The Region of Southern Denmark.
